THESIS PROJECT PROPOSAL

### 2019-11-11 project_ashish/paper.tex
Update and rename paper1.tex to paper.tex

\begin{figure}
  \centering
   \includegraphics[scale=0.5]{/home/ashish/tHqCMSAnalysis/project_ashish/cd7.png}
  \caption{ATLAS-CMS combined measurements of coupling strength modifiers.}
   \label{figure 3}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{/home/ashish/tHqCMSAnalysis/project_ashish/cms5.png}
  \caption{Standard model Higgs boson production in association with top quark and top anti-quark, top quark decaying to W boson and bottom quark and  W boson decaying to leptons.}
   \label{figure 4}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{/home/ashish/tHqCMSAnalysis/project_ashish/cms13.png}
  \caption{Scan of $−2Δln(L)$ versus $κ_t$ for the data (black line) and the individual channels (blue, red, and green), compared to Asimov data sets corresponding to the SM expectations (dashed lines).}
   \label{figure 5}
\end{figure}

Until the 90's the existence of almost all the particles of the SM were confirmed except the top quark and the Higgs boson. 
These had eluded previous experiments due to difficulties in the production or reconstruction of its decays. The top quark was 
discovered in 1995 in the Tevatron collider of the Fermilab laboratory, this proton collider operated with a center of mass 
energy of 1.8 TeV untill 2010. The LHC collider at the CERN laboratory in Geneva, Switzerland, began its operations in 2010 
colliding protons at 7 TeV increasing the colliding energies to 8 and 13 TeV in the subsequent years. The Compact Muon Solenoid (CMS) is based on the Large Hadron Collider (LHC) located at CERN. It is designed to detect particles known as muons very accurately. The CMS detector has the form of a cylindrical onion, with several concentric layers of components. Needed a powerful magnet to bend charged particles as they move away from the point of collision to identify the charge of the particles to bend them in opposite directions and measure your momentum. A silicon tracker, made of about 75 
million electronic sensors individual arranged in concentric layers, identifies the routes taken by these charged particles bent with very high precision \cite{Chatrchyan:2008aa}. The muons are not detected by electromagnetic or hadron calorimeters, so special sub-detectors are outside to detect them once they have crossed the solenoid as shown in Figure 6. In 2012, the ATLAS and CMS collaborations, with detectors at two points where the protons collide in the LHC, announced the discovery of a new boson with a mass of 125 GeV. So far, all measurements of the properties of this boson are consistent with those of the Higgs boson of the standard model (SM). LHC during the first run in 2011 and 2012 reached a peak luminosity of 7.7 \times 10^{33} $cm^{-2}s^{-1}$ which was more than 75$\%$ of its design luminosity and delivered an integrated luminosity of about 25 $fb^{-1}$ to each ATLAS and CMS. The goal of subsequently run of LHC has been to obtain datasets at higher values of luminosities for precision measurements of the properties of Higgs boson, in order to test the Standard Model pattern of couplings to elementary particles. In order to minimize the machine downtimes and maximize the productive use of the LHC for physics, the replacement of the inner triplet magnets (the one responsible to squeeze the beam at collision) and  of  all  hardware  changes  needed  to  enable  an  ambitious  luminosity  upgrade  will  take  placein parallel during one shutdown, at around 2023-25 (Long shutdown 3), with some of the modification anticipated in 2019-2020 (LS2). This new phase of the LHC life has been named as High Luminosity LHC (HL-LHC) and has the scope of attaining the astonishing threshold of 3000 $fb^{-1}$ in 10-12 years. All the hadron colliders in the world have so far produced a total integrated luminosity of about 10 $fb^{-1}$, while the LHC will deliver about 300 $fb^{-1}$ in its first 10-12 years of life \cite{collaborations2019report}. LHC will reach an integrated luminosity of about 300 $fb^{-1}$ by 2023 as shown in Figure 7. \\

\begin{figure}
  \centering
   \includegraphics[scale=0.5]{/home/ashish/tHqCMSAnalysis/project_ashish/cms12.png}
  \caption{Transverse view of the CMS detector showing the silicon tracker, superconducting solenoid, electromagnetic calorimeter, hadron calorimeter and muon chambers.}
   \label{figure 6}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{/home/ashish/tHqCMSAnalysis/project_ashish/lum6.png}
  \caption{\onehalfspacing Projected performance of the LHC until 2038, which shows the preliminary dates for prolonged stops (LS) of the LHC and luminosities. Points
show instant brightness while the line shows brightness accumulated.}
   \label{figure 7}
\end{figure}

The production of Higgs boson in association with single top quark is one of the rare Higgs 
boson production mode \cite{Khachatryan_2016}. As the top 
quark is the heaviest fundamental particle in the standard model, due to its large mass, the 
top quark decay before 
hardronization which allow the possibility to reconstruct top quark from its decay products 
unlike the lighter quarks which 
undergo hardronization and are seen as bundles of particles in detector called jets. The most 
probable decay of top quark is 
into bottom (b) quark and W boson \cite{nishiwaki2014tth}. Using b-jet tagging algorithms, the jet originating from 
b-quark can be reconstructed to 
identity the top quark. In SM, the coupling of the Higgs boson to fermions like quarks and 
leptons is proportional to the mass 
of the fermions. Thus heavy quarks like top, bottom and charm couple strongly to Higgs boson 
which means that out of all known 
quarks, top quark couple most strongly to the Higgs boson. So, it has a large value of Yukawa 
coupling $y_t$ \cite{Aad_2016}. That is why the 
Yukawa coupling of the top quark with the Higgs boson $y_t$ has lot of importance as any 
deviation from standard model prediction might give an indication of new physics. The two 
processes at LHC which will allow us to directly probe the Yukawa 
coupling between Higgs Boson and top quark are Higgs boson production in association with top 
quark pair (t$\bar{t}$H) via 
strong interaction and the production of single top quark with a Higgs boson (tH). As tH 
occur via electroweak interaction, it 
is more rare than t$\bar{t}$H production \cite{sirunyan2018search}. The Feynman diagrams for SM Higgs boson in association with top quark and top anti-quark is shown in Figure 8. The Higgs boson in the tH channel can be radiated 
off by a W boson or by a top 
quark as shown in Figure 8 and 9. These two processes in standard model have destructive interference which allow us to 
probe the relative sign between 
the coupling of top quark with Higgs boson $y_t$ and the coupling of W boson to the Higgs 
boson \cite{choudhury2019search}. If the Yukawa coupling between top quark and Higgs boson $y_t$ deviate from the SM 
prediction or even if $y_t$ has a negative sign, both would cause a strong increase in the tH 
production cross section which is a special property of tH production channel making it an 
interesting production channel to probe using LHC 2016, 2017 and 2018 data. The coupling 
strength modifier $\kappa_t$ is defined as $\kappa_t = \frac{y_t}{y^{SM}_t}$ which will tell 
us the deviation of $y_t$ from the SM prediction as shown in Figure 5. 
The LHC is sensitive to an anamolous Higgs coupling to the top quark in the Higgs boson-top 
quark associated production mode. 
The anomalous coupling will arise when we add interaction terms (dimension 6 operators in 
effective SM theory) in the Higgs 
potential. As there is strong destructive inteference in the t-channel for the standard model 
couplings, this production mode 
is sensitive to both the sign and the magnitude of any coupling beyond the SM between Higgs 
boson and top quark induced by 
higher dimensional operators in effective SM theory \cite{sirunyan2019search}.

\begin{figure}
  \centering
   \includegraphics[scale=0.5]{/home/ashish/tHqCMSAnalysis/project_ashish/tq.png}
  \caption{Higgs production using the top quark (left) and W boson (right).}
  \label{figure 8}

  \begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/cms.png}
  \caption{Leading order Feynman diagrams for the associated production of a single top quark and a Higgs boson in the t channel, where the Higgs boson couples either to the top quark (left) or the W boson (right).}
   \label{figure 9}
\end{figure}

@@

\newpage
\cleanpage

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/cms7.jpg}
  \caption{Higgs boson signal strength modifier in various production modes and decay channels.}
   \label{figure 9}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/cd.png}
  \caption{Leading order Feynman Diagrams for various Higgs production modes.}
   \label{figure 10}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/cd1.png}
  \caption{Standard model Higgs boson production cross section in various production modes.}
   \label{figure 11}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/cd2.png}
  \caption{Standard model Higgs boson production cross section with center of mass energy and branching ratios for various decay channels.}
   \label{figure 12}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[scale=0.5]{/home/ashish/tHqCMSAnalysis/project_ashish/cd3.png}
  \caption{Standard model Higgs boson branching ratios in various decay channels.}
   \label{figure 13}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/cd5.png}
  \caption{Combined measurement of the products $\sigma$.BR for various production and decay channels of Standard model Higgs boson.}
   \label{figure 14}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/cd6.png}
  \caption{Likelihood contours in the ($\kappa_F$,$\kappa_V$) plane and combined fit for all decay channels.}
   \label{figure 15}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/cd7.png}
  \caption{ATLAS-CMS combined measurements of coupling modifiers.}
   \label{figure 16}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/th.png}
  \caption{Feynman diagrams contributing to the partonic process W b $->$ tH.}
   \label{figure 17}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/th1.png}
  \caption{Partonic cross sections for the process W b $->$ tH as a function of the center of mass
energy.}
   \label{figure 18}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/th2.png}
  \caption{Lowest order Feynman diagrams for t-channel tH production in the 4F scheme and in the 5F scheme.}
   \label{figure 19}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{/home/ashish/tHqCMSAnalysis/project_ashish/th3.png}
  \caption{Lowest order Feynman diagrams for s-channel tH production.}
   \label{figure 20}
\end{figure}

>>>>

The production of Higgs boson in association with single top quark is one of the rare Higgs 
boson production mode \cite{Khachatryan_2016}. As the top 
quark is the heaviest fundamental particle in the standard model, due to its large mass, the 
top quark decay before 
hardronization which allow the possibility to reconstruct top quark from its decay products 
unlike the lighter quarks which 
undergo hardronization and are seen as bundles of particles in detector called jets. The most 
probable decay of top quark is 
into bottom (b) quark and W boson \cite{nishiwaki2014tth}. Using b-jet tagging algorithms, the jet originating from 
b-quark can be reconstructed to 
identity the top quark. In SM, the coupling of the Higgs boson to fermions like quarks and 
leptons is proportional to the mass 
of the fermions. Thus heavy quarks like top, bottom and charm couple strongly to Higgs boson 
which means that out of all known 
quarks, top quark couple most strongly to the Higgs boson. So, it has a large value of Yukawa 
coupling $y_t$ \cite{Aad_2016}. That is why the 
Yukawa coupling of the top quark with the Higgs boson $y_t$ has lot of importance as any 
deviation from standard model prediction might give an indication of new physics. The two 
processes at LHC which will allow us to directly probe the Yukawa 
coupling between Higgs Boson and top quark are Higgs boson production in association with top 
quark pair (t$\bar{t}$H) via 
strong interaction and the production of single top quark with a Higgs boson (tH). As tH 
occur via electroweak interaction, it 
is more rare than t$\bar{t}$H production \cite{sirunyan2018search}. The Feynman diagrams for SM Higgs boson in association with top quark and top anti-quark is shown in Figure 3.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{./cms5.png}
	\caption{\onehalfspacing Standard model Higgs boson production in association with top quark and top anti-quark, top quark decaying to W boson and bottom quark and  W boson decaying to leptons.}
	\label{figure 3}
\end{figure}

 The Higgs boson in the tH channel can be radiated 
off by a W boson or by a top 
quark as shown in Figure 4 and 5.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{./tq.png}
	\caption{Higgs boson production using the top quark (left) and W boson (right).}
	\label{figure 4}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\columnwidth]{./cms.png}
	\caption{\onehalfspacing Leading order Feynman diagrams for the associated production of a single top quark and a Higgs boson in the t channel, where the Higgs boson couples either to the top quark (left) or the W boson (right).}
	\label{figure 5}
\end{figure}

 These two processes in standard model have destructive interference which allow us to 
probe the relative sign between 
the coupling of top quark with Higgs boson $y_t$ and the coupling of W boson to the Higgs 
boson \cite{choudhury2019search}. If the Yukawa coupling between top quark and Higgs boson $y_t$ deviate from the SM 
prediction or even if $y_t$ has a negative sign, both would cause a strong increase in the tH 
production cross section which is a special property of tH production channel making it an 
interesting production channel to probe using LHC 2016, 2017 and 2018 data \cite{farina2013lifting}. The coupling 
strength modifier $\kappa_t$ is defined as $\kappa_t = \frac{y_t}{y^{SM}_t}$ which will tell 
us the deviation of $y_t$ from the SM prediction as shown in Figure 6.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{./cd7.png}
	\caption{ATLAS-CMS combined measurements of coupling strength modifiers.}
	\label{figure 6}
\end{figure}

The LHC is sensitive to an anamolous Higgs coupling to the top quark in the Higgs boson-top 
quark associated production mode. 
The anomalous coupling will arise when we add interaction terms (dimension 6 operators in 
effective SM theory) in the Higgs 
potential. As there is strong destructive inteference in the t-channel for the standard model 
couplings, this production mode 
is sensitive to both the sign and the magnitude of any coupling beyond the SM between Higgs 
boson and top quark induced by 
higher dimensional operators in effective SM theory \cite{sirunyan2019search}.

Until the 90's the existence of almost all the particles of the SM were confirmed except the top quark and the Higgs boson. 
These had eluded previous experiments due to difficulties in the production or reconstruction of its decays. The top quark was 
discovered in 1995 in the Tevatron collider of the Fermilab laboratory, this proton collider operated with a center of mass 
energy of 1.8 TeV untill 2010. The LHC collider at the CERN laboratory in Geneva, Switzerland, began its operations in 2010 
colliding protons at 7 TeV increasing the colliding energies to 8 and 13 TeV in the subsequent years. The Compact Muon Solenoid (CMS) is based on the Large Hadron Collider (LHC) located at CERN. It is designed to detect particles known as muons very accurately. The CMS detector has the form of a cylindrical onion, with several concentric layers of components. Needed a powerful magnet to bend charged particles as they move away from the point of collision to identify the charge of the particles to bend them in opposite directions and measure your momentum. A silicon tracker, made of about 75 
million electronic sensors individual arranged in concentric layers, identifies the routes taken by these charged particles bent with very high precision \cite{Chatrchyan:2008aa}. The muons are not detected by electromagnetic or hadron calorimeters, so special sub-detectors are outside to detect them once they have crossed the solenoid as shown in Figure 7. In 2012, the ATLAS and CMS collaborations, with detectors at two points where the protons collide in the LHC, announced the discovery of a new boson with a mass of 125 GeV. So far, all measurements of the properties of this boson are consistent with those of the Higgs boson of the standard model (SM).

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.36]{./cms12.png}
	\caption{Transverse view of the CMS detector showing the silicon tracker, superconducting solenoid, electromagnetic calorimeter, hadron calorimeter and muon chambers.}
	\label{figure 7}
\end{figure}

 LHC during the first run in 2011 and 2012 reached a peak luminosity of 7.7 $\times$ $10^{33}$ $cm^{-2}s^{-1}$ which was more than 75$\%$ of its design luminosity and delivered an integrated luminosity of about 25 $fb^{-1}$ to each ATLAS and CMS. The goal of subsequently run of LHC has been to obtain datasets at higher values of luminosities for precision measurements of the properties of Higgs boson, in order to test the Standard Model pattern of couplings to elementary particles. In order to minimize the machine downtimes and maximize the productive use of the LHC for physics, the replacement of the inner triplet magnets (the one responsible to squeeze the beam at collision) and  of  all  hardware  changes  needed  to  enable  an  ambitious  luminosity  upgrade  will  take  placein parallel during one shutdown, at around 2023-25 (Long shutdown 3), with some of the modification anticipated in 2019-2020 (LS2). This new phase of the LHC life has been named as High Luminosity LHC (HL-LHC) and has the scope of attaining the astonishing threshold of 3000 $fb^{-1}$ in 10-12 years. All the hadron colliders in the world have so far produced a total integrated luminosity of about 10 $fb^{-1}$, while the LHC will deliver about 300 $fb^{-1}$ in its first 10-12 years of life \cite{collaborations2019report}. LHC will reach an integrated luminosity of about 300 $fb^{-1}$ by 2023 as shown in Figure 8. \\

 \begin{figure}[H]
 	\centering
 	\includegraphics[scale=0.45]{./lum6.png}
 	\caption{\onehalfspacing Projected performance of the LHC until 2038, which shows the preliminary dates for prolonged stops (LS) of the LHC and luminosities. Points show instant brightness while the line shows brightness accumulated.}
 	\label{figure 8}
 \end{figure}

 Earlier searches for Higgs boson production production in association with single top quark have been studied to derive constraints on the magnitude and relative sign of Higgs boson couplings to top quarks and vector bosons. Constaints on $\kappa_t$ can be derived using a likelihood ratio scan of $L(\kappa_t)/L(\hat{\kappa_t})$ where $\hat{\kappa_t}$ is the best fit value of $\kappa_t$ as shown in Figure 9. Multilepton final state and final states with single lepton with a pair of bottom quarks were combined with Higgs decay to two photons to get the final result. The data favour positive value of top quark yukawa coupling for SM Higgs boson and excludes ranges of about [-0.9,-0.5] and [1.0,2.1] times $y^{SM}_t$ at the 95 $\%$ confidence level.

 \begin{figure}[H]
  \centering
   \includegraphics[width=\columnwidth]{./cms13.png}
  \caption{Scan of -2 $\Delta$ ln(L) versus $\kappa_t$ for the data (black line) and the individual channels (blue, red, and green), compared to Asimov data sets corresponding to the SM expectations (dashed lines).}
   \label{figure 9}
\end{figure}

@@

\newpage

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./cms7.jpg}
  \caption{Higgs boson signal strength modifier in various production modes and decay channels.}
   \label{figure 10}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./cd.png}
  \caption{Leading order Feynman Diagrams for various Higgs production modes.}
   \label{figure 11}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./cd1.png}
  \caption{Standard model Higgs boson production cross section in various production modes.}
   \label{figure 12}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./cd2.png}
  \caption{Standard model Higgs boson production cross section with center of mass energy and branching ratios for various decay channels.}
   \label{figure 13}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[scale=0.5]{./cd3.png}
  \caption{Standard model Higgs boson branching ratios in various decay channels.}
   \label{figure 14}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./cd5.png}
  \caption{Combined measurement of the products $\sigma$.BR for various production and decay channels of Standard model Higgs boson.}
   \label{figure 15}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./cd6.png}
  \caption{Likelihood contours in the ($\kappa_F$,$\kappa_V$) plane and combined fit for all decay channels.}
   \label{figure 16}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./th.png}
  \caption{Feynman diagrams contributing to the partonic process W b $->$ tH.}
   \label{figure 17}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./th1.png}
  \caption{Partonic cross sections for the process W b $->$ tH as a function of the center of mass
energy.}
   \label{figure 18}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./th2.png}
  \caption{Lowest order Feynman diagrams for t-channel tH production in the 4F scheme and in the 5F scheme.}
   \label{figure 19}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{./th3.png}
  \caption{Lowest order Feynman diagrams for s-channel tH production.}
   \label{figure 20}
\end{figure}
